---
title: "Ileostomy Site Carcinoma Manuscript"
author: "Anthony Morada"
date: "12/14/2020"
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
  word_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

# Loading Data
```{r setup, echo = FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(kableExtra)
library(readxl)
library(dplyr)
library(janitor)
library(table1)
library(finalfit)
library(tableone)


df <- read_excel("~/THE GUTHRIE CLINIC/Foundation Research Projects - Ileostomy Site Carcinoma/case table.xlsx", na = "NA") 
df <- df %>%
  clean_names() %>%
  filter(!is.na(exclusion))
```
The table has `r ncol(df)` columns and `r nrow(df)` rows.  

# Full-Text Inclusion and Exclusion 
```{r inclusion/exclusion}
df_i <- df[df$exclusion == "No",]
df_e <- df[df$exclusion == "Yes",]

ex_cit <- subset(df_e, !(df_e$citation %in% df_i$citation)) %>%
  select(citation, exclusion_reason) %>%
  unique()
```

Our table contains `r length(unique(df$citation))` unique papers.  

We included `r nrow(df_i)` patients that were found in `r length(unique(df_i$citation))` unique citations. We excluded `r nrow(ex_cit)` citations. Below is reason for full-text exclusions:  
```{r printing exclusion reason summary}
summary(as.factor(ex_cit$exclusion_reason))
```

# Cleaning Included Manuscripts

```{r setting working data frame}
df <- df_i
rm(df_i, df_e, ex_cit)

# Removing columns from editors
df <- select(df, -checked_by_2_people, -comments, -chai, -harsha, -anthony, -amir, -exclusion, -exclusion_reason, -time_from_index_pathology_years, -patient_number)

```

__Creating Column for Malignant or Benign Etiology:__  
  
* Malignant Etiology = Adenocarcinoma, Lymphoma, Melanoma, Neuroendocrine Tumor, Squamous Cell Carcinoma, Verrucous Carcinoma  
* Pre-Malignant Etiology = ESFA, Extramammary Paget Disease  
  
```{r cleaning}
# Creating Malignant etiology column
df$malignant <- grepl("Adenocarcinoma|Lymphoma|Melanoma|Neuroendocrine Tumor|Squamous Cell Carcinoma|Verrucous Carcinoma", df$final_pathology, ignore.case = T)
df$malignant <- gsub(TRUE, "Malignant", df$malignant)
df$malignant <- gsub(FALSE, "Pre-malignant", df$malignant)
df$malignant <- as.factor(df$malignant)

# Creating ibd-fap column
df <- df %>%
  mutate(ibd_fap = case_when(index_surgery_indication == "Ulcerative Colitis" ~ "IBD",
                             index_surgery_indication == "Crohn's Disease" ~ "IBD",
                             index_surgery_indication == "FAP" ~ "FAP",
                             TRUE ~ "Other"))

# Creating binary lymph, binary mets, and binary recurrence/mets column
df <- df %>% 
  mutate(binary_lymph = case_when(simple_lymph == "1" ~ "Positive",
                                  simple_lymph == "3+" ~ "Positive",
                                  is.na(simple_lymph) ~ NA_character_,
                                  TRUE ~ "Negative"
                                  )) %>%
  mutate(binary_mets = case_when(metastasis == "Negative" ~ "Negative",
                                 metastasis == "Positive" ~ "Positive",
                                 is.na(metastasis) ~ NA_character_,
                                 TRUE ~ "Positive")) %>%
  mutate(binary_recurr = case_when(recurrance_mets_later == "none" ~ "Negative",
                                   is.na(recurrance_mets_later) ~ NA_character_,
                                   TRUE ~ "Positive"
                                   )) %>%
  mutate(ac_scc = case_when(final_pathology == "Adenocarcinoma" ~ "Adenocarcinoma",
                             final_pathology == "Squamous Cell Carcinoma" ~ "Squamous Cell Carcinoma",
                             is.na(final_pathology) ~ NA_character_,
                             TRUE ~ "Other"
                             ))

# Making all character columns into factors
df[sapply(df, is.character)] <- lapply(df[sapply(df, is.character)],  as.factor)

#For TableOne
kableone <- function(x, ...) {
  capture.output(x <- print(x))
  knitr::kable(x, ...)
}

```


# Data Analysis  

## Table 1  
```{r table 1}
#Variables for analysis
vars = c("age_at_index_surgery", "age_of_ileostomy_site_neoplasm_diagnosis", "time_from_index_surgery_age_of_stoma", "sex", "race", "index_surgery_indication", "ibd_fap", "largest_dimension_cm", "biopsy_simplified", "time_between_initial_and_excisional_biopsies_year", "final_pathology", "ac_scc", "malignant" ,"binary_lymph","binary_mets","resection","fulguration","chemo","radiation","binary_recurr","survival_period_years","status")

# Locating columns with non normal distribution
cont_vars <- colnames(df[,sapply(df, function(x) is.numeric(x))])
sapply(df[,cont_vars],shapiro.test)
print(sapply(df[,cont_vars],shapiro.test))
nonnorm_vars <- c("age_at_index_surgery","largest_dimension_cm","time_between_initial_and_excisional_biopsies_year", "survival_period_years")
print(paste("Column with a non-normal distribution:", nonnorm_vars))

#Print Table One
kableone(print(
  CreateTableOne(vars = vars, data = df), 
  nonnormal = nonnorm_vars,
  showAllLevels = TRUE,
  missing = TRUE
  ))

```


## Pre-Malignant vs Malignant  
```{r}
stratify <- "malignant"

#Fisher's exact test
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[cat_vars %in% vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]][[1]],df[stratify][[1]]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify), 
  nonnormal = nonnorm_vars,
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
```

## Adenocarcinoma vs Squamous  
```{r}
stratify <- "ac_scc"

#Fisher's exact test
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[cat_vars %in% vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]][[1]],df[stratify][[1]]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify), 
  nonnormal = nonnorm_vars,
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
```

## FAP vs IBD  
```{r}
stratify <- "ibd_fap"

#Fisher's exact test
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[vars %in% cat_vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]][[1]],df[stratify][[1]]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify), 
  nonnormal = nonnorm_vars,
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
```

## Mets and Recurrence
```{r}
stratify <- "binary_recurr"

#Fisher's exact test
fischer_var = c()
cat_vars <- colnames(df[,sapply(df, function(x) is.factor(x))])
cat_vars <- cat_vars[vars %in% cat_vars]
for (i in 1:length(cat_vars)) {
  if(sum(chisq.test(table(df[,cat_vars[i]][[1]],df[stratify][[1]]))$expected < 5) > 0) {
    fischer_var[i] <- cat_vars[i]
  }
}

kableone(print(
  CreateTableOne(vars = vars, data = df, strata = stratify), 
  nonnormal = nonnorm_vars,
  exact = fischer_var,
  showAllLevels = TRUE,
  missing = TRUE
  ))
```

## Presentation Frequency
```{r}
present <- unique(as.character(df$simple_presentation))
present <- present[!is.na(present)]
present <- paste(present, collapse = "; ")
present <- strsplit(present, "; ")
present <- present[[1]]
present <- unique(present)
```

```{r}
for(i in 1:length(present)){
     print("--------------------------------------------------------------------------")
     print(present[i])
     print(sum(grepl(present[i],df$simple_presentation)))
     print(table(grepl(present[i], df$simple_presentation), df$ibd_fap))
     print(fisher.test(table(grepl(present[i], df$simple_presentation), df$ibd_fap))$p.value)
     print(table(grepl(present[i], df$simple_presentation), df$ac_scc))
     print(fisher.test(table(grepl(present[i], df$simple_presentation), df$ac_scc))$p.value)
}
```
## Age of stoma  
```{r}

library(ggplot2)
mu <- plyr::ddply(df, "ibd_fap", summarise, grp.mean=mean(time_from_index_surgery_age_of_stoma,na.rm = TRUE))
mu <- filter(mu, !mu$ibd_fap == "Other")

df %>%
 filter(!is.na(time_from_index_surgery_age_of_stoma)) %>%
 filter(!(ibd_fap %in% 
    "Other")) %>%
 ggplot() +
 aes(x = time_from_index_surgery_age_of_stoma, fill = ac_scc) +
 geom_histogram(bins = 30L) +
 scale_fill_grey() +
 geom_vline(data=mu, aes(xintercept=grp.mean, linetype=ibd_fap)) + 
 labs(x = "Stoma Age (Years)", y = "Count", fill = "Ileostomy Pathology", linetype = "Initial Surgery Indication") +
 theme_gray() +
 facet_grid(vars(ibd_fap), vars())

ggsave("age_of_stoma.tiff", 
       width = 19.05, 
       height = 22.23, 
       units = "cm",
       dpi = 300
       )

```



```{r}
## Screening


length((df[df$ibd_fap == "IBD",]$time_from_index_surgery_age_of_stoma >=25 | df[df$ibd_fap == "IBD",]$age_of_ileostomy_site_neoplasm_diagnosis >= 60) & 
  (!is.na(df[df$ibd_fap == "IBD",]$time_from_index_surgery_age_of_stoma) & !is.na(df[df$ibd_fap == "IBD",]$age_of_ileostomy_site_neoplasm_diagnosis)))

sum((df[df$ibd_fap == "IBD",]$time_from_index_surgery_age_of_stoma >=25 | df[df$ibd_fap == "IBD",]$age_of_ileostomy_site_neoplasm_diagnosis >= 60) & 
  (!is.na(df[df$ibd_fap == "IBD",]$time_from_index_surgery_age_of_stoma) & !is.na(df[df$ibd_fap == "IBD",]$age_of_ileostomy_site_neoplasm_diagnosis)))


length((df[df$ibd_fap == "FAP",]$time_from_index_surgery_age_of_stoma >=20 | df[df$ibd_fap == "FAP",]$age_of_ileostomy_site_neoplasm_diagnosis >= 50) & 
  (!is.na(df[df$ibd_fap == "FAP",]$time_from_index_surgery_age_of_stoma) & !is.na(df[df$ibd_fap == "FAP",]$age_of_ileostomy_site_neoplasm_diagnosis)))

sum((df[df$ibd_fap == "FAP",]$time_from_index_surgery_age_of_stoma >=20 | df[df$ibd_fap == "FAP",]$age_of_ileostomy_site_neoplasm_diagnosis >= 50) & 
  (!is.na(df[df$ibd_fap == "FAP",]$time_from_index_surgery_age_of_stoma) & !is.na(df[df$ibd_fap == "FAP",]$age_of_ileostomy_site_neoplasm_diagnosis)))

```