---
title: "Ileostomy Site Carcinoma Literature Review"
author: "Anthony Morada"
date: "11/11/2020"
output:
  html_document:
    toc: yes
    df_print: paged
  word_document:
    toc: yes
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

## Loading and Cleaning

```{r setup, echo = FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE
)

library(readxl)
library(dplyr)
library(janitor)
library(table1)
library(finalfit)

df <- read_excel("~/THE GUTHRIE CLINIC/Foundation Research Projects - Ileostomy Site Carcinoma/case table.xlsx", na = "NA") 
df <- df %>%
  clean_names() %>%
  filter(!is.na(exclusion))

```
The table has `r ncol(df)` columns and `r nrow(df)` rows.

## Full-Text Inclusion and Exclusion 

```{r inclusion/exclusion}
df_i <- df[df$exclusion == "No",]
df_e <- df[df$exclusion == "Yes",]

ex_cit <- subset(df_e, !(df_e$citation %in% df_i$citation)) %>%
  select(citation, exclusion_reason) %>%
  unique()
```

Our table contains `r length(unique(df$citation))` unique papers 

We included `r nrow(df_i)` patients that were found in `r length(unique(df_i$citation))` unique citations. We excluded `r nrow(ex_cit)` citations. Below is reason for full-text exclusions:  
```{r printing exclusion reason summary}
summary(as.factor(ex_cit$exclusion_reason))
```

## Summary of Data  

Malignant Etiology = Adenocarcinoma, Lymphoma, Melanoma, Neuroendocrine Tumor, Squamous Cell Carcinoma, Verrucous Carcinoma  
Pre-Malignant Etiology = ESFA, Extramammary Paget Disease  
  
```{r creating new column}
df_i$malignant <- grepl("Adenocarcinoma|Lymphoma|Melanoma|Neuroendocrine Tumor|Squamous Cell Carcinoma|Verrucous Carcinoma", df_i$final_pathology, ignore.case = T)
df_i$malignant <- gsub(TRUE, "Malignant", df_i$malignant)
df_i$malignant <- gsub(FALSE, "Pre-malignant", df_i$malignant)
df_i$malignant <- as.factor(df_i$malignant)

df_i <- df_i %>%
  mutate(ibd_fap = case_when(index_surgery_indication == "Ulcerative Colitis" ~ "IBD",
                             index_surgery_indication == "Crohn's Disease" ~ "IBD",
                             index_surgery_indication == "FAP" ~ "FAP",
                             TRUE ~ "Other"))
```




```{r table 1}
library(table1)
table1::label(df_i$age_of_ileostomy_site_neoplasm_diagnosis) <- "Age of Ileostomy Site Malignancy Diagnosis"
table1::label(df_i$sex) <- "Sex"
table1::label(df_i$race) <- "Race"
table1::label(df_i$time_from_index_surgery_age_of_stoma) <- "Age of Stoma"
table1::label(df_i$final_pathology) <- "Pathology"
table1::label(df_i$malignant) <- "Pathology Potential"
table1::label(df_i$index_surgery_indication) <- "Index Surgery Indication"
table1::label(df_i$ibd_fap) <- "Generalized Surgery Indication"
table1::label(df_i$survival_period_years) <- "Follow Up (Years)"
table1::label(df_i$status) <- "Follow Up Status"


table1::table1(~ age_of_ileostomy_site_neoplasm_diagnosis + 
                 sex +
                 race +
                 time_from_index_surgery_age_of_stoma +
                 final_pathology +
                 malignant +
                 index_surgery_indication +
                 ibd_fap + 
                 survival_period_years +
                 status 
               ,data = df_i)

```

### Pre-Malignant vs Malignant  
```{r benign vs malignant}
dependent = "malignant" 
explanatory = c("age_of_ileostomy_site_neoplasm_diagnosis", "sex", "race","time_from_index_surgery_age_of_stoma", "final_pathology","index_surgery_indication", "ibd_fap")

table <- df_i %>%
  summary_factorlist(dependent, explanatory,
                     p=TRUE, na_include = TRUE)

knitr::kable(table, align=c("l", "l", "r", "r", "r"))
```

### Adenocarcinoma vs Squamous  
```{r}
dependent = "final_pathology" 
explanatory = c("age_of_ileostomy_site_neoplasm_diagnosis", "sex", "race","time_from_index_surgery_age_of_stoma", "index_surgery_indication", "ibd_fap")

table <- df_i %>%
  filter(final_pathology == "Adenocarcinoma" | final_pathology == "Squamous Cell Carcinoma") %>%
  summary_factorlist(dependent, explanatory,
                     p=TRUE, na_include = TRUE)

knitr::kable(table, align=c("l", "l", "r", "r", "r"))
```

### FAP vs IBD  
```{r}
dependent = "ibd_fap" 
explanatory = c("age_of_ileostomy_site_neoplasm_diagnosis", "sex", "race","time_from_index_surgery_age_of_stoma", "final_pathology", "malignant", "index_surgery_indication")
table <- df_i %>%
  summary_factorlist(dependent, explanatory,
                     p=TRUE, na_include = TRUE)

knitr::kable(table, align=c("l", "l", "r", "r", "r"))
```



